### 2013-03-22 document.tex
Signed-off-by: loopasam <samuel.croset@gmail.com>

pouet \cite{Krotzsch2012} pouet

The introduction should put the focus of the manuscript into a broader context. 
As you compose the introduction, think of readers who are not experts in this field. 
Include a brief review of the key literature. If there are relevant controversies or disagreements in 
the field, they should be mentioned so that a non-expert reader can delve into these issues further. 
The introduction should conclude with a brief statement of the overall aim of the experiments and a comment 
about whether that aim was achieved.

>>>>

The introduction \cite{Krotzsch2012} should put the focus of the manuscript into a broader context. 
As you compose the introduction, think of readers who are not experts in this field. 
Include a brief review of the key literature. If there are relevant controversies or disagreements in 
the field, they should be mentioned so that a non-expert reader can delve into these issues further. 
The introduction should conclude with a brief statement of the overall aim of the experiments and a comment 
about whether that aim was achieved.

Functional Therapeutic Chemical Classification System (FTC) is a taxonomy (ontology) representing the mode of action of approved
drugs in the human body.
It can be pictured as a toolbox classifying the therapeutic agents based on their functionality in the human body.
FTC categories are automatically built based on Gene Ontology terms.
FTC categories are also equivalent to arbitrary definitions such as drug that perturbs some (protein that involved-in some (regulation of 
biological process)).
A knowledge base created by integrating the content of DrugBank and some GO annotations assists to automatically
assign compounds into FTC categories based on the satisfaction of the definition.
The FTC is evaluated against the Anatomical Therapeutic Chemical Classification System by assigning equivalences between classes.
The knowledge base can be queried in order to retrieve implicit information about the involvement of approved drugs into biological
processes.
Drug repurposing hypotheses can be formulated out of the evaluation.
The FTC can serve as starting point for analysis or projects requiring a structured vocabulary representing the mode of action.
The source code of the FTC as well as the data are publicly available for free to anyone.

The indication of a drug depends on the effects induced by the molecule when administrated in a biological system.
For instance, a molecule such as the Ximelagatran is indicated as anticoagulant because
it directly inhibits the catalytic activity of the thrombin, 
enzyme itself involved in the blood coagulation.
In order to speak of this therapeutic effect (anticoagulant), people sometimes refer to 
the **mode of action** of an active compound. 
Wikipedia gives us the following definition for it:
> A mode of action describes a functional or anatomical change, at the 
> cellular level, resulting from the exposure of a living organism to a substance.

Characterising the mode of action of drugs is interesting for the following reasons:
It is a central link between the molecular structure of a drug and its indication towards a disease. Drugs with similar
mode of actions could be used to treat similar conditions.
It abstracts away from the chemical structure and allows to reason over the effects of the compound in the body. The mode of action
of a drug is a feature against which compounds can be classified.
It can captures polypharmacology. Most of the drugs are not magic bullets
and produce various side effects. A drug can have several mode of actions, representing the different processes the drug is perturbing directly
or indirectly.
A representation of the mode of action helps to better understand and analyse the behaviour of an agent in the human body.

Although the **mode of action** is an old and useful notion in the drug discovery 
community (see number of results for thisPubMed search, no resources were
previously fully dedicated to it. The aim of the FTC is to fill this gap.

The FTC can be seen as a toolbox designed to fix dysfunctioning biological systems.
It features over 20'000 compartments (categories) inside which approved 
drugs are automatically classified, based on facts from various databases.
FTC categories represent the mode of action of compounds in the human system. 
The resource can be used for various tasks where a structured vocabulary is needed, such as compound annotation, drug repurposing analysis and 
anything else you can think of!
